# Missing values
```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(readr)
library(tidyverse)
library(socviz)
library(dplyr)
library("ggpubr")
library(openintro)
library(dbplyr)
library(car)
library(highcharter)
library(funModeling)
library(patchwork)
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv',show_col_types = FALSE)
songs_info <- spotify_songs %>%
  select(-c(track_id,track_name,track_artist,track_album_id,track_album_name,track_album_release_date,playlist_name,playlist_id,playlist_subgenre))
```

```{r,fig.width=20,fig.height=20,echo=FALSE,message=FALSE,warning=FALSE}
missing_value_plot<-function(data,type='counts'){
  # find and count the missing patterns
  missing_patterns <- data.frame(is.na(data)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup()
  
  # add id to each row
  missing_patterns <- missing_patterns %>% 
    mutate(id=row_number(),.before=1)
  
  # tidy the missing patterns
  new_missing_patterns <- missing_patterns %>%
    mutate_if(is.logical, as.character)
    #mutate_all(funs(as.character(.)))
  for (i in 1:nrow(new_missing_patterns)) {
    flag <-0
    for (j in 2:ncol(new_missing_patterns)) {
      if (new_missing_patterns[i,j]=='TRUE') {
        flag <- 1
      } 
    }
    if (flag==0) {
      new_missing_patterns[i,2:(ncol(new_missing_patterns)-1)] <-'complete'
    }
  }
  tidypatterns <- new_missing_patterns %>%
    select(-count) %>%
    pivot_longer(cols=!id,names_to ='key',values_to='value')
  
  # Plot the main graph
  column_sum <- colSums(is.na(data))
  variable_order <- names(sort(column_sum,decreasing = TRUE))
  variable <- fct_relevel(tidypatterns$key,variable_order)
  main <- ggplot(tidypatterns,aes(x=variable,y=fct_rev(factor(id)),fill=value))+
    geom_tile(color = "white")+
    scale_fill_manual(values = c('TRUE'='darkorchid1','FALSE'='grey','complete'='darkgrey'))+
    theme(legend.position = "none")+
    labs(x='variable',y='missing pattern')
  
  # add text
  complete_loc <- tidypatterns$id[which(tidypatterns$value=='complete')]
  x <- as.integer(ncol(data)/2)+1
  for (i in complete_loc) {
    y <- nrow(missing_patterns)-i+1
    main <- main+geom_text(aes(x=x,y=y,label='complete case'),size=4,family="Times New Roman",fontface='bold')
  }
  
  if (type=='counts'){
    # plot the top graph
    column_count<-data %>% 
          summarise_all(~ sum(is.na(.x))) %>%
          pivot_longer(cols=everything(), names_to="variable", values_to="count") %>% 
          arrange(desc(count))
    top <- ggplot(column_count,aes(x=factor(variable, levels=column_count$variable), y=count))+
      geom_col(fill='lightblue')+
      labs(x='',y='num rows missing:')+
      ggtitle('Missing Value Patterns with counts')+ theme_grey(16) 
    
    # plot the right graph
    row_count <-new_missing_patterns %>%
      add_column(status='missing')
    for (i in 1:nrow(row_count)) {
      if (any(row_count[i,]=='complete')){
        row_count[i,'status']<-'complete'
      }
    }
    right <- ggplot(row_count,aes(x=count,y=fct_rev(factor(id)),fill=status))+
      geom_col()+
      scale_fill_manual(values=c('blue','lightblue'))+
      theme(legend.position = "none")+
      labs(x='row count',y='')
  }
  else if (type=='percents'){
    # plot the top graph
    column_count <- data %>%
      summarise_all(~ 100*sum(is.na(.x))/nrow(data)) %>% 
      pivot_longer(cols=everything(), names_to="variable", values_to="count") %>% 
      arrange(desc(count))
    top <- ggplot(column_count,aes(x=factor(variable, levels=column_count$variable), y=count))+
      geom_col(fill='lightblue')+
      scale_y_continuous(limits=c(0,100),breaks=seq(0,100,by=25))+
      labs(x='',y='% rows missing:')+
      ggtitle('Missing Value Patterns with percents')
    
    # plot the right graph
    row_count <-new_missing_patterns %>%
      add_column(status='missing') %>%
      mutate(count=100*(count/nrow(data))) %>%
      arrange(desc(count))
    for (i in 1:nrow(row_count)) {
      if (any(row_count[i,]=='complete')){
        row_count[i,'status']<-'complete'
      }
    }
    right <- ggplot(row_count,aes(x=count,y=fct_rev(factor(id)),fill=status))+
      geom_col()+
      scale_fill_manual(values=c('blue','lightblue'))+
      scale_x_continuous(limits=c(0,100),breaks=seq(0,100,by=25))+
      theme(legend.position = "none")+
      labs(x='% rows',y='')
  }
  else (print('Wrong Type Input'))
  
  top + plot_spacer() + main + right + 
    plot_layout(
      ncol = 2, 
      nrow = 2, 
      widths = c(4, 1),
      heights = c(1, 4)
    )
}
```

```{r,fig.width=26,fig.height=26,echo=FALSE,message=FALSE,warning=FALSE}
missing_value_plot(spotify_songs[0:11],type = 'counts')
```

```{r,fig.width=20,fig.height=20,echo=FALSE,message=FALSE,warning=FALSE}
missing_value_plot(spotify_songs[12:23],type = 'counts')
```
<br>
Because our dataset has too many columns and each column name is long, it's not easy to see the missing pattern of the whole data in one graph, so instead, we separated the dataframe into 2 parts and plot each part. As we can see from the graph that in our dataset, there are only three columns with missing values, `track_artist`,`track_name`, and `track_album_name`. Since all three columns contains information that is not relevant to our analysis on music characteristic and genre and we will not use those three columns at all, so we decided to omit it.